---
title: "Lab exercise: day 4"
editor: visual
---

```{r}
#| label: setup

library(tsibble)
library(tsibbledata)
library(fable)
library(feasts)
library(fabletools)
library(tidyverse)
```

# Learn

Produce forecasts from an automatically selected ETS model for Australia's print media turnover. Does the chosen ETS model align with the patterns you see in the data?

```{r}
aus_print <- aus_retail |> 
  filter(Industry == "Newspaper and book retailing") |> 
  summarise(Turnover = sum(Turnover))

fit <- aus_print |> 
  model(
    ???
  )
fit
fit |> 
  ???(h = "10 years") |> 
  autoplot(???)
```

Produce forecasts from an automatically selected ARIMA model for Australia's print media turnover. Don't forget to find and use a suitable transformation to simplify the multiplicative patterns in the data.

```{r}
# Find an appropriate transformation
aus_print |> 
  autoplot(Turnover)

# Estimate an ARIMA model
fit <- aus_print |> 
  model(
    ???
  )
fit
fit |> 
  ???(h = "10 years") |> 
  autoplot(???)
```

# Apply

```{r}
vaccine_administrated_tsb <- read_rds("data/vaccine_administrated_tsb.rds")
```

## Specify and traing ETS models

We now want to apply the Exponential Smoothing family of models to forecast dose administrated.

> Remember the following from ETS (E: Error, T:Trend, S: Seasonality) function:
>
> -   N: None (No trend, no seasonality)
> -   A: additive
> -   Ad: additive damped
> -   M: multiplicative

Complete the R code to train exponential smoothing models on the `vaccine_administrated_tsb`:

```{r}
fit_ets <- vaccine_administrated_tsb |>
  model(
    ??? = ???(???)
  )
```

> If you don't provide terms inside ETS(), then it is an automatic ETS! It will examine different models and return the one with lowest AICc.

## Observe model table and extract model's output

let's now observe the model table `fit_ets`. What type of data structure is `fit_ets`? How many rows and columns are present, and what do they represent?

Complete the below R codes and run to understand model's outputs:

Use `report()`:

```{r}
fit_ets |> filter(region == "") |> ???()
```

Which model is returned in, what are its components and parameter values?

Use `tidy()`:

```{r}
fit_ets |> ???()
```

Use `glance()`:

```{r}
fit_ets |> ???()
```

You can also observe the values corresponding to level, trend,and seasonal components in ETS framework. Each column corresponds to one components.

Complete the R code and run to observe the selected model components:

```{r}
fit_ets |> ???()
```

Could you describe what each row and column represent?

## Generate forecasts using ETS

Now, we can forecast with the trained ETS model. Complete and run the following code to produce forecasts:

```{r}
forecast_horizon <- 12
forecast_ets <- fit_ets |>
  ???(h = ???) 
forecast_ets
```

What type of data structure is it? What each row and column represent?

## Visualise forecasts:

You can also visualize forecasts:

```{r}
forecast_ets |>
  autoplot(filter_index(???, "2020" ~ .), level = NULL)
```

## Detrmine model components and parameters manually

You can use the following specific functions: `error()`, `trend()`, `season()` to manually specify the type of pattern ("N", "A", "Ad", "M") and also their corresponding parameters, if you wish.

In the following R-chunk, you can change parameters and the type of pattern to see their impact on fitting and forecast:

```{r alpha}
vaccine_ets <- vaccine_administrated_tsb |> filter(region == "") |> 
  model(
    `alpha = 0.05` = ETS(??? ~ error("A") + trend("A", alpha = .05) + season("N")),
    `alpha = 0.15` = ETS(??? ~ error("A") + trend("A", alpha = .15) + season("N")),
    `alpha = 0.5` = ETS(??? ~ error("A") + trend("A", alpha = .5) + season("N")),
    `alpha = 0.9` = ETS(??? ~ error("A") + trend("A", alpha = .9) + season("N"))
  )

vaccine_ets |> augment() |> filter_index("2020" ~ .) |> 
  ggplot(aes(x=month))+
  geom_line(aes(y=dose_adminstrated, colour= "Actual"))+
  geom_line(aes(y=.fitted, colour= factor(.model)))+
  ggthemes::scale_color_colorblind()+
  labs(colour ="")
```

Could you describe how the value of smoothing constant , alpha affects the forecast?

## ARIMA

## Specify and train ARIMA model

> If you want this function automatically determines the order of autoregressive and moving average orders and their parameters, then you don't need to provide arguments inside ARIMA(). The function will examine different models (combinations of p=0,1,2,.. and q =0,1,2,.., P=0,1,2,.. and Q =0,1,2,..) and return the one with lowest AICc.

Complete the following code to train an automatic ARIMA model on the data:

```{r}
fit_arima <- vaccine_administrated_tsb |>
  model(
    ??? = ???(???)
  )
```

> If we don't provide terms inside ARIMA(), then it is an automatic model! It will examine different models and return the one with lowest AICc.

let's now observe the fitted model, `fit_arima`:

```{r}
fit_arima
```

What type of data structure is it? What each row and column represent?

## Extract model's output

use `report()`

```{r}
fit_arima |> filter(region == "") |> 
  ???()
```

Which model is returned in, what are its components and parameter values?

Use `tidy()`:

```{r}
fit_arima |> ???()
```

Use `glance()`:

```{r}
fit_arima |> ???()
```

## Generate forecasts using ARIMA

Now, we can forecast with the fitted ARIMA model:

```{r}
forecast_horizon <- 12
vaccine_fcst_arima <- ??? |> ???(h = ???)  
```

Let's now observe the forecast table, `vaccine_fcst_arima`:

```{r}
vaccine_fcst_arima
```

What type of data structure is it? What each row and column represent?

## Visualize forecasts

You can visualize your forecast:

```{r}
vaccine_fcst_arima |> 
  autoplot(filter_index(???, "2020" ~ .))
```

## Determine model components manually

You can also specify the order of p,q,P,Q using specific function `pdq()` and `PDQ()` manually, complete the r code

```{r}
fit_arima_manual <- ??? |>
  model(
    automatic_arima = ???(???),
    arima1=???(??? ~ ???(1,1,1)+???(0,0,1)),
    arima2 = ???(??? ~ ???(3,1,0)+???(1,0,0))
  )
fit_arima_manual
```

Could you check AICc for the manually determined models and see how it compares to the automatic selected ARIMA model?

```{r}
fit_arima_manual |> glance() |> select(region,.model, AICc) |>  arrange(region,AICc)
```
